\OnePageSectionStart
\section{Formal Semantics and Graph Theory}

\subsection{Denotational Semantics of Typed Tensors}
A typed tensor is interpreted as a pair: value and semantic grounding.

\[ \llbracket \text{Tensor}\langle D, S, \Sigma \rangle \rrbracket = (v, \sigma) \]

where $v \in \mathbb{R}^{D}$ (the value) and $\sigma \in \text{SemanticSpace}$ (the meaning).

\textbf{Semantic spaces}:
\begin{itemize}
  \item $\text{Atomic}: \emptyset$ (no semantic constraint)
  \item $\text{Measurement}: (\text{PhysicalQuantity}, \text{Unit}, \text{Frame}) \times [a, b]$ (grounded in physical reality with bounds)
  \item $\text{Distribution}: (\text{Entity}, \text{StochasticProcess}, \text{Support})$ (probabilistic model of an entity)
  \item $\text{Enum}: \text{ConceptScheme} \times \mathbb{N}$ (symbolic vocabulary with cardinality)
  \item $\text{Structured}: \prod_i \text{SemanticSpace}_i$ (product of semantic spaces)
\end{itemize}

A morph $M$ is a morphism that preserves semantic structure:
\[ P: (v_1, \sigma_1) \times \cdots \times (v_n, \sigma_n) \to (v'_1, \sigma'_1) \times \cdots \times (v'_m, \sigma'_m) \]

\textbf{Semantic preservation constraint}: $P$ must satisfy $\sigma'_j \sqsubseteq \sigma_j$ (semantic subsumption) or involve an explicit type bridge.

\subsection{Graph-Theoretic Model of Computation}
A computation graph $G = (N, E, \tau)$ where:
\begin{itemize}
  \item $N$ is a set of morph nodes
  \item $E \subseteq N \times N$ are directed edges (data flow)
  \item $\tau: E \to \text{Type}$ assigns a type to each edge
\end{itemize}

The graph must satisfy:
\begin{enumerate}
  \item \textbf{Well-typedness}: for each edge $(u, v) \in E$, output type of $u$ is compatible with input type of $v$
  \item \textbf{Acyclicity}: no cycles (data flow is a DAG)
  \item \textbf{Semantic closure}: semantic types form a lattice under subsumption; each edge type is a lattice element
\end{enumerate}

\subsection{Three Orthogonal Graph Layers}

A statically typed system actually operates over three intertwined graphs:

\paragraph{1. Computation Graph}
The data flow DAG: $G_{\text{comp}} = (N_{\text{prim}}, E_{\text{data}})$ where edges carry tensor types. This is what executes at runtime.

\paragraph{2. Type Graph}
The constraint system over types: $G_{\text{type}} = (N_{\text{type}}, E_{\text{subtype}})$ where nodes are types and edges are subsumption relations. The system checks that every data edge respects type constraints.

\paragraph{3. Semantic Graph}
The RDF/ontology layer: $G_{\text{sem}} = (N_{\text{iri}}, E_{\text{predicate}})$ where nodes are IRIs (physical quantities, knowledge concepts, stochastic processes) and edges are RDF predicates. This ground-truth graph describes what meanings the computation is over.

\textbf{Invariant}: For every edge $e \in E_{\text{data}}$ with type $\tau(e)$, the semantics of $\tau(e)$ must resolve via $G_{\text{sem}}$. If $\tau(e) = \text{Measurement}\{\text{velocity}, \text{unit:m/s}\}$, then $\text{velocity}$ and $\text{m/s}$ must be IRIs resolvable in $G_{\text{sem}}$.

\subsection{The Fabric of Reality, Knowledge, and Computation}

These three graphs form a unified fabric:
\begin{itemize}
  \item **Reality Fabric** ($G_{\text{sem}}$ layer): the ontology of physical quantities, measured by sensors
  \item **Knowledge Fabric** ($G_{\text{sem}}$ layer): the vocabulary of symbols and concepts (tokens, entities, relations)
  \item **Computation Fabric** ($G_{\text{comp}}$ layer): the neural network DAG that transforms observations
  \item **Type Fabric** ($G_{\text{type}}$ layer): the constraints ensuring safe composition
\end{itemize}

These fabrics are not separate---they are woven together. A Measurement tensor carries both:
- A value (computation fabric): the pixel values from a camera
- A semantic IRI (reality fabric): the physics quantity ``radiance'' with unit ``sRGB''

A Distribution tensor carries:
- A value (computation fabric): probabilities summing to 1
- A semantic IRI (knowledge fabric): the decision space it models

When morphs compose, all three fabrics must align. This is why semantic typing is powerful: it enforces alignment across all three simultaneously.

\OnePageSectionEnd
