\OnePageSectionStart
\section{Graph-Cognitive Fusion: The Dynamics of $\CognitiveFabric$}

Reasoning is the composition of functional blocks. In the Cognitive Fabric $\CognitiveFabric$, we treat this composition as a formal \emph{graph fusion} operation.

\begin{axiom}[Structural Fusion]
The fusion of two compute graphs $G_1$ and $G_2$ across a set of ports $P = \{ (in_i, out_j) \}$ is well-defined if and only if the semantic types of $in_i$ and $out_j$ are unifyable under $\RealityFabric$ subsumption.
\end{axiom}

\subsection{Cognitive Morphs as Graph Envelopes}
While individual nodes (Add, Mul, Conv) are the "atoms" of the machine, we define \emph{Cognitive Morphs} as higher-level envelopes that hide computational complexity. A Morph is a graph that implements a specific reasoning pattern (e.g., "Spatial Reasoning", "Linguistic Inference").

\subsection{Deterministic Composition via RDF CONSTRUCT}
A key operational observation is that RDF/SPARQL can serve not only for querying but as a deterministic composition language. SPARQL's \texttt{CONSTRUCT} form allows the specification of graph-transformation templates that map ontology-aligned motifs and praxis into fused compute graphs. This gives humans and machines a shared, machine-readable script for composition: a \texttt{CONSTRUCT} query is a deterministic, auditable program that derives new triples (a fused graph) from existing ontology and instance data.

Practically, composition flows look like:
\begin{enumerate}
  \item Use ontology-driven SPARQL to select motifs and compatible ports (based on semantic subsumption and constraints).
  \item Apply \texttt{CONSTRUCT} templates to assemble a fused RDF description of the target cognitive fabric (nodes, edges, semantics, pragmas).
  \item Lower the constructed RDF graph into Fuse/ModelProto artifacts and validate with SHACL and the compiler's type checks.
\end{enumerate}

This deterministic composition is "very meta": it allows a machine intelligence (for example, a coding LLM that produces Fuse EBNF snippets in a single-shot) to synthesize candidate graphs, query the ontology to verify compatibility, and then emit a valid cognitive fabric. Because the entire pipeline is RDF-driven, it is transparent and auditable: each constructed triple has provenance and can be debugged by humans or further processed by reasoners.

\OnePageSectionEnd
